E exp 3

### 2018-08-25 experiments/D_exp_2/writeup/D_exp_2.Rmd
think i forgot to commit some old changes

The instructions for this experiment were similar to C_exp_2, except that we changed "fewer than" to "far fewer than": this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.

@@

```{r}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
```

@@

```{r}
dat$Condition <- as.factor(paste(dat$Item, dat$Vagueness))
ggplot(data=dat, mapping=aes(x=Item, y=RT_log, group=Vagueness)) +
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), aspect.ratio = 1) +
  facet_wrap(~Selection) +
  stat_summary(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), fun.y=mean, geom="line", lwd=2) +
  geom_smooth(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), method="loess")
```

@@

Testing for a main effect of X when X is a two-level factor and the model also includes a main effect of a three-level factor  and the interaction between X and Y.

Testing for a main effect of Vagueness when Vagueness is a two-level factor and the model also includes a main effect of a four-level factor Item and the interaction between Vagueness and Item.

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]
m0 <- lm(RT_log ~ 
           Item1 + Vagueness:Item1 + 
           Item2 + Vagueness:Item2 + 
           Item3 + Vagueness:Item3 +
           Selection +
           Selection:Vagueness, 
         dat)
contrasts(dat$Item) <- contr.sum(4)
m.test <- lm(RT_log ~ 
               Item +
           Vagueness:Item1 + 
           Vagueness:Item2 + 
           Vagueness:Item3 +
           Selection +
           Selection:Vagueness, 
         dat)
m1 <- lm(RT_log ~ 
           Item1 + Vagueness:Item1 + 
           Item2 + Vagueness:Item2 + 
           Item3 + Vagueness:Item3 +
           Selection +
           Selection:Vagueness +
           Vagueness,
         dat)
m100 <- lm(RT_log ~
             Item +
             Vagueness +
             Item:Vagueness+
             Selection+
             Selection:Vagueness,
           dat)
#m1 and m100 are the same model
anova(m0,m1,test="F")
```

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]
m3 <- lm(RT_log ~ 
           Item1 + 
           Item2 +  
           Item3 , 
         dat)
contrasts(dat$Item) <- contr.sum(4)
m4 <- lm(RT_log ~
           Item,
         dat)
# m3 and m4 are the same model
```

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]

m.full <- lm(RT_log ~
             Item +
             Vagueness +
             Selection + 
             Item:Vagueness +
             Item:Selection + 
             Vagueness:Selection,
           dat)
m.reduced <- lm(RT_log ~
                Item1 + 
                Item2 + 
                Item3 + 
                Vagueness:Item1 + 
                Vagueness:Item2 +
                Vagueness:Item3 +
                Vagueness:Selection +
                Selection +
                Item:Selection
                ,dat  )
anova(m.reduced, m.full, test="F")
```

# Discussion

>>>>

We changed "fewer than" (crisp) to "far fewer than" (in vague): this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.

@@

```{r, cho=FALSE}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
```

@@

@@

The variables are:
Vagueness 2 levels (X) 
Item 4 levels (Y) 
Selection 2 levels (Z)

Note that the null-hypothsesis model _includes_ random slopes for the main effect of X; Y; and Z; as well as for their interactions -- including a seperate random slope for all combinations of X, Y, and Z achieves this.

```{r}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))

Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
dat_model$Item1 <- Item.numeric[1,]
dat_model$Item2 <- Item.numeric[2,]
dat_model$Item3 <- Item.numeric[3,]
```

```{r}
null.model2 <- lmer(RT_log ~ 
                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
                   dat_model, REML=FALSE) 
```

```{r}
pretty_coef_table(null.model2, "null.model2")
```

```{r}
full.model2 <- lmer(RT_log ~ 
                     c_Vag + 
                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
                   dat_model, REML=FALSE) 
```

```{r}
pretty_coef_table(full.model2, "full.model2")
```

```{r}
anova(null.model2, full.model2)
```

* [Hypothesis 1] Crisp/Vague RT:
    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions on average (beta = -0.0057 units of log(ms)). However this difference was not significant in the full model (p=0.678). Using Levy's method to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.6764) - consituting more evidence that Vagueness did not exert a significant main effect on response times. 
* [Hypothesis 2] Comparison/Matching RT:
    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (beta=0.1618, p<0.001).
* [Hypothesis 3] Vagueness effect differs between comparison and matching
    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant (beta=0.1206, p<0.001), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task. Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).

### 2018-08-25 experiments/E_exp_3/writeup/E_exp_3.Rmd
think i forgot to commit some old changes

```{r, "make data for rt plot"}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
```

@@

The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)

# Discussion

>>>>

```{r, "make data for rt plot", echo=FALSE}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
```

@@

The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018).

The variables are:
Vagueness 2 levels (X) 
Item 4 levels (Y) 
Selection 2 levels (Z)

Note that the null-hypothsesis model _includes_ random slopes for the main effect of X; Y; and Z; as well as for their interactions -- including a seperate random slope for all combinations of X, Y, and Z achieves this.

```{r}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))

Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
dat_model$Item1 <- Item.numeric[1,]
dat_model$Item2 <- Item.numeric[2,]
dat_model$Item3 <- Item.numeric[3,]
```

```{r}
null.model2 <- lmer(RT_log ~ 
                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
                   dat_model, REML=FALSE) 
```

```{r}
pretty_coef_table(null.model2, "null.model2")
```

```{r}
full.model2 <- lmer(RT_log ~ 
                     c_Vag + 
                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
                   dat_model, REML=FALSE) 
```

```{r}
pretty_coef_table(full.model2, "full.model2")
```

```{r}
anova(null.model2, full.model2)
```

* [Hypothesis 1] Crisp/Vague RT:
    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions on average (beta = -0.015 units of log(ms)). However this difference was not significant in the full model (p=0.119). Using Levy's method to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.118) - consituting more evidence that Vagueness did not exert a significant main effect on response times.

* [Hypothesis 2] Comparison/Matching RT:
    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (beta=0.176, p<0.001).

* [Hypothesis 3] Vagueness effect differs between comparison and matching:
    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : 
    Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant (beta=0.123, p<0.001), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task. Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).
